---
title: "Propensity Score Implementation"
author: "Grayson Ricketts"
date: "11/7/2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(dev = 'pdf', fig.path = 'Figs/',
                      fig.width = 8.5, fig.height = 5,
                      echo = TRUE, warnings = FALSE)

library(ggplot2)
theme_set(theme_minimal())

library("faraway")
library("RItools")
library("gbm")
library("optmatch")
```

```{r, echo=FALSE}
lalonde <- read.table("job_training_observational_data.txt", header = TRUE)

# Removes redundant variable already represented by TREAT
lalonde <- within(lalonde, rm("TYPE"))

# Keep dataset with RE78 values
lalonde.full <- lalonde

# Removes the response variable so that the design is not biased by outcomes
lalonde <- within(lalonde, rm("RE78"))
```

# Logistic Regression of Propensity Scores

## Standard Model
```{r}
model <- glm(TREAT ~ . , family = binomial(link = "logit"), data = lalonde)
summary(model)
```

## Reduced Model
```{r}
# Reduces model based on non-significant predictors in full model
bad.factors <- c("NODEGREE", "EDUC", "RE74")
lalonde.reduced <- lalonde[ , !(names(lalonde) %in% bad.factors)]

model.reduced <- glm(TREAT ~ . , family = binomial(link = "logit"), data = lalonde.reduced)
summary(model.reduced)
```

## Model Checking

### Half-Normal Plot of Residuals
```{r}
halfnorm(residuals(model))
halfnorm(residuals(model.reduced))
```

### Chi-Squared Goodness-of-Fit-Test for Model
```{r}
sum(residuals(model, type = "pearson")^2)
deviance(model)
pchisq(deviance(model), df.residual(model))
pchisq(deviance(model.reduced), df.residual(model.reduced))

xBalance(TREAT ~ . - (TREAT) , 
  data = lalonde.reduced, 
  report = c("chisquare.test"))
```

### Hosmer-Lemeshow Test
```{r}
hosmerlem <- function (y, yhat, g = 10) {
   cutyhat <- cut(yhat, breaks = quantile(yhat, probs = seq(0, 1, 1/g)),
                  include.lowest = TRUE)
   obs <- xtabs(cbind(1 - y, y) ~ cutyhat)
   expect <- xtabs(cbind(1 - yhat, yhat) ~ cutyhat)
   chisq <- sum((obs - expect)^2 / expect)
   P <- 1 - pchisq(chisq, g - 2)
   c("X^2" = chisq, Df = g - 2, "P(>Chi)" = P)
}
hosmerlem(y = lalonde$TREAT, yhat = fitted(model))
hosmerlem(y = lalonde$TREAT, yhat = fitted(model.reduced))
```

## Summary of Model Checking
It appears that both models are accurate in their prediction. This can be seen in the low Chi-Squared probabilities of the standard Chi-Squared Goodness-of-Fit-Test and the Hosmer-Lemeshow test.

## Using Estimated Propensity Scores to Create Predicted Values
```{r}
# Attach propensity score value to full dataset
lalonde.full$propscore.log <- predict(model, type = "response")
lalonde.reduced$propscore.log <- predict(model.reduced, type = "response")

ggplot(lalonde.full[lalonde.full$RE74 < quantile(lalonde.full$RE74, .98), ], 
       aes(propscore.log, fill = factor(TREAT))) +
  geom_histogram(aes(y = ..density..), alpha = .5, position = "identity")

ggplot(lalonde.reduced[lalonde.reduced$RE75 < quantile(lalonde.reduced$RE75, .98), ], 
       aes(propscore.log, fill = factor(TREAT))) +
  geom_histogram(aes(y = ..density..), alpha = .5, position = "identity")
```

# Generalized Boosted Model (GBM) Propensity Scores

```{r}
# Suggested parameters comes from Guo, 2001
model.generalized.suggested <- gbm(TREAT ~ ., distribution = "bernoulli", data = lalonde, 
                         train.fraction = 0.5, 
                         interaction.depth = 4, 
                         shrinkage = 0.0005)
model.generalized.default <- gbm(TREAT ~ ., distribution = "bernoulli", data = lalonde)

summary(model.generalized.suggested)
summary(model.generalized.default)

lalonde.full$propscore.gbm <- predict(model.generalized.suggested, type = "response")
```

# Matching
```{r}
match.pair <- pairmatch(mdist(model.reduced), data = lalonde.reduced)
match.full <- fullmatch(mdist(model.reduced), data = lalonde.reduced)
```

## Balance Testing
```{r}
(allbalance <- xBalance(TREAT ~ . -(TREAT) , 
    data = lalonde.reduced, 
    report = c("chisquare.test", "std.diffs"), 
    strata = data.frame(original = factor("none"), match.reduced, match.full)))
```

Full match has a p-value of 0.917 which is perfect. It appears that the imbalances have been sorted out.

# Analysis

## Weighted Analysis
Weights for ATE
```{r}
lalonde.reduced$weight.ATE <- ifelse(lalonde.reduced$TREAT == 1, 1 / lalonde.reduced$propscore.log, 
                                     1 / (1 - lalonde.reduced$propscore.log))
lalonde.full$gweight.ATE <- ifelse(lalonde.full$TREAT == 1, 1 / lalonde.full$propscore.gbm, 
                                     1 / (1 - lalonde.full$propscore.gbm))
```

Check balance of weights
```{r}
summary(lm(RE75 ~ TREAT, data = lalonde.reduced, weights = lalonde.reduced$weight.ATE))
summary(lm(RE75 ~ TREAT, data = lalonde.full, weights = lalonde.full$gweight.ATE))
```

Balancing using wieghts seems to be really bad.



## Matched Analysis

## Naive Analysis

